---
title: "Maps of Erica"
author: "Duncan Sephton"
date: "2025-03-03"
output: html_document
---
#Packages used:
```{r}
library(sf)
library(tidyverse)
library(lwgeom)
library(terra)
library(stars)
library(exactextractr)
library(ggplot2)
library(rinat)
library(ggspatial)
library(prettymapr)
library(leaflet)
library(htmltools)
library(wesanderson)
library(mapview)
library(leafpop)
```
# Reading in the data for soil map of South Africa:
```{r}
veg <- st_read("SOTER/ZA_SOTERv1.shp")
veg               
st_crs(veg)
class(veg)
head(veg)

# Plotting the map
ggplot() +
  geom_sf(data=veg, aes(fill = `SOIL`))
```






# Cropping the data to the extent of the area I know the species that I'll use are located in:
```{r}
myextent <- st_sf(a = 1:2, geom = st_sfc(st_point(c(19,-33.5)), st_point(c(23,-35))), crs = 4326)
sf_use_s2(FALSE)

vet <- st_crop(veg, myextent)

```
Plotting Cropped Data
```{r}
# plotting the cropped data for soil data
ggplot() +
  geom_sf(data=vet, aes(fill = `SOIL`))
```




# Data for Erica discolor:
### Erica discolor data pulled in from iNat
```{r}
# collecting iNat data
eos <- get_inat_obs(taxon_name = "Erica discolor",
                   bounds = c(-35, 19, -33.5, 22),
                   maxresults = 1000)

eot <- st_as_sf(eos, coords = c("longitude", "latitude"), crs = 4326)

# mapping raw data
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data=eot)

#Filtering 
eo <- eos %>% filter(positional_accuracy<46 & 
                      latitude<0 &
                      !is.na(latitude) &
                      captive_cultivated == "false" &
                      quality_grade == "research")
eo <- st_as_sf(eo, coords = c("longitude", "latitude"), crs = 4326)


#Plotting the filtered data
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data=eo)


```





#Mapping Erica oraria:
### Erica oraria pulled in from iNat
```{r}
# collecting iNat data
ep <- get_inat_obs(taxon_name = "Erica oraria",
                   bounds = c(-35, 19, -33.5, 22),
                   maxresults = 1000)
# mapping the raw data

ept <- st_as_sf(ep, coords = c("longitude", "latitude"), crs = 4326)

ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data=ept)

# Filtering iNat data for only research grade observations
eps <- ep %>% filter(positional_accuracy<46 & 
                latitude<0 &
                !is.na(latitude) &
                captive_cultivated == "false" &
                quality_grade == "research")

# mapping with research grade observations

eps <- st_as_sf(eps, coords = c("longitude", "latitude"), crs = 4326)

ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data=eps)

```





# Connecting the iNat data to the iNat site with a URL while also mapping for both Inat species:
```{r}
# Linking the url for Inat
eor <- eo %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))

# Erica discolor map
eot <- mapview(eo,
        layer.name = c("Erica discolor"),
        
        col.regions = c("#7680BB"), 
        popup = 
          popupTable(eor,
            zcol = c("scientific_name", "click_url")))
# Linking the url for Inat
epts <- eps %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))

# Erica oraria map
ept <- mapview(eps,
        layer.name = c("Erica oraria"),
        
        col.regions = c("#F8805D"),
        popup = 
          popupTable(epts,
            zcol = c("scientific_name", "click_url")))

# Combing maps
mapv <- ept + eot 


```
# Making an Interactive Soil map:
```{r}

pal <- wes_palette("Darjeeling1", 9, type = "continuous")

Bio <- mapview(vet, 
              zcol = ("SOIL"),
             popup = 
          popupTable(vet,
            zcol = c("SOIL")))
```
# Combining soil and iNat maps to make the final map:
```{r}
Bio + mapv
```







